---
title: "excel_password_accessing"
author: "Dave Lovell"
date: "14/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reading Password Protected Excel Files

## TLDR

If you're after a minimal explanation with a function you can copy and paste, scroll straight to the very end of this document.

## Setup

There is an Excel file in this directory that's protected by a password

```{r files}
"file_with_a_password.xlsx" -> xl_file
```

Obviously we can't read it with `readxl::read_xlsx`:

```{r dontWork}
try((readxl::read_xlsx(xl_file)))
```

## Reading the File with RDCOM

I found this solution [here](https://stat.ethz.ch/pipermail/r-help/2011-March/273678.html).

RDCOMClient is a package that interfaces with a lot of Microsoft stuff. It's **non-CRAN** and the current build (`omegahat/RDCOMClient`) is **incompatible with Rtools 4.0**. For this reason, we'll need to use `dkyleward/RDCOMClient`, which has been fixed by a helpful person on the internet. 

```{r useRDCOM}
# devtools::install_github("dkyleward/RDCOMClient")
library("RDCOMClient")
```

## Make an eAPP
We need to create an object in R that can interact with excel files (or something). We do this by specifying class `Excel.Application` in the `RDCOMClient::COMCreate`. Note that this won't work on a machine that doesn't have Excel installed.

```{r createApp}
eApp <-  COMCreate("Excel.Application")
```

## Open the File

The object we've just created is odd: we can 'COM Object Methods' on it (from it?) with the `$` operator. Doing so returns S4 objects of the same class as our `eAPP`. In the code below, we call the `Workbooks()` and `Open()` to open a workbook, specifying the `Filename` and `Password` arguments. Note that these arguments start with capital letters. 

```{r readFile}
# RDCOM appears not to like relative filenames
xl_file <- glue::glue("{getwd()}/{xl_file}")

unlocked_file <-  eApp$Workbooks()$Open(
  Filename = as.character(xl_file), # Open() can't handle `glue` class.
  Password = "top secret password") # I use this password for most of my accounts, so don't share it with anyone.
```
You'll might notice a coule of oddities in the above code:

* I've used `glue` and `getwd()` to get the full name of the file, all the way back to the home directory. RDCOMClient appears not understand filenames that are relative to the wd.
* I've subsequently coerced the result of `glue()` into a character vector. `class(xl_file)` gives `r class(xl_file)`, and the `glue` class causes the function to throw an error. This package is obviously very sensitive. 

## Write File Contents to Table

To write the file contents to a table, use `$Sheets()$SaveAs()`, specifying the sheet in the prior function and the filename (and the number three) in the latter. I would just use a tempfile for this. 

```{r getContents}
# Create a temporary .txt file
tf <- tempfile(fileext = ".txt")

## Save the contents of sheet one to the tempfile. Note that we're not creating an Excel file here.
unlocked_file$Sheets(1)$SaveAs(tf, 3) # I'm not sure what '3' is for, but don't change it.

out_table <- read.table(tf, sep = "\t")

out_table
```

You'll likely want to do a bit of `colnames(out_table) <- out_table[1,]; out_table <- out_table[-1,]`. If you want to create an non-password protected excel file, you'll need `xlsx::write.xlsx(as.data.frame(out.table))`, but bear in mind that you may lose some formatting / metadata in conversion. RDCOMClient probably has a more direct way of doing that, but I don't know what it is. 

### Package excel.link

There's a package called 'excel.link' which puts a nice function wrapper around all of this, but that package uses the old build of RDCOMClient, which causes R to crash. Furthermore, loading 'excel.link' will cause any RDCOMClient code to crash, because it will install and attach the old version over the top of the one that works. Took me a while to work that out!

```{r CRANMethod}
### DO NOT RUN THIS ----------------
# library("excel.link")

# excel.link::xl.read.file(file = "file_with_a_password.xlsx",
#                            password = "top secret password",
#                            write.res.password = "top secret password",
#                            xl.sheet = "Sheet1") -> excelOut
```

## Creating a Fucntion (and TLDR)

### TLDR

Run `devtools::install_github("dkyleward/RDCOMClient")`. It's important that you install this version specifically. 

Steal this function:

### Function

```{r excelReaderFunction}
read_protected_excel <- function(filename = NULL, password = NULL, sheet = 1, root = getwd(),
                                 as.tibble = TRUE, fix.col.names = TRUE) {
  stopifnot(is.character(filename))
  stopifnot(is.character(password))
  stopifnot(is.numeric(sheet))
  
  eApp <-  RDCOMClient::COMCreate("Excel.Application")
  
  xl_file <- as.character( glue::glue("{root}/{filename}") )
  password <- as.character(password)

unlocked_file <-  eApp$Workbooks()$Open(
  Filename = as.character(xl_file), 
  Password = password)

  tf <- tempfile(fileext = ".txt")
  unlocked_file$Sheets(1)$SaveAs(tf, 3)
  out_table <- read.table(tf, sep = "\t")

  if(fix.col.names) names(out_table) <- out_table[1,]; out_table <- out_table[-1,]
  if(as.tibble) out_table <- (tibble::as_tibble(out_table))
  
  return(out_table)
}

read_protected_excel(filename = "file_with_a_password.xlsx", 
                     password = "top secret password")
```
