---
title: "excel_password_accessing"
author: "Dave Lovell"
date: "14/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Excel Files

```{r files}
dir(pattern = ".*\\.xlsx")
```

One of them is password protected, one isn't.

## Copying Some Code From the Internet

[Here](https://stat.ethz.ch/pipermail/r-help/2011-March/273678.html), specifically.

You'll need to install package `RDCOMClient`, which is non-CRAN, apparently.

```{r internetCode}
# # devtools::install_github("omegahat/RDCOMClient")
# library("RDCOMClient")
# eApp <-  COMCreate("Excel.Application.14")
# wk <-  eApp$Workbooks()$Open(Filename="your_file",Password="your_password")
# tf <-  tempfile()
# wk$Sheets(1)$SaveAs(tf, 3)
```

Okay, struggling to get the above to work because my personal laptop doesn't have Excel and my work laptop doesn't have Rtools! Need to get the latter sorted. I think I might have found a CRAN method though, looks way simpler (but I suspect is using RDCOM under the hood?)

## CRAN Method

```{r CRANMethod}
library("excel.link")

excel.link::xl.read.file(file = "file_with_a_password.xlsx",
                         password = "top_secret_password",
                         # write.res.password = "top_secret_password",
                         xl.sheet = "Sheet1") -> excelOut
```

The above code crashed R a couple of times, I'm not sure why. 